---
title: "STA6235 Week 2"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
  error: true
editor: source
---

## Tuesday, August 29

1.  Pull in the [Palmer penguins dataset](https://allisonhorst.github.io/palmerpenguins/).

```{r}
install.packages("tidyverse")
install.packages("dplyr")
install.packages("palmerpenguins")
library(palmerpenguins)
library(tidyverse)
library(dplyr)
data(package = 'palmerpenguins')
head(penguins, n=100)
print(penguins)
```

2.  Find the mean and standard deviation for the body mass of penguins (*body_mass_g*). Please use the `summarize()` function from `tidyverse`.

```{r}
penguins %>%
summarize(mean=mean(body_mass_g, na.rm=TRUE),
          std_dev=sd(body_mass_g, na.rm=TRUE))
```

3.  Find the mean and standard deviation for the body mass of penguins (*body_mass_g*), split by species of penguin (*species*). Please use the `group_by()` and `summarize()` functions from `tidyverse`.

```{r}
penguins %>%
   group_by(species) %>% 
summarize(mean=mean(body_mass_g, na.rm=TRUE),
          std_dev=sd(body_mass_g, na.rm=TRUE))
```

4.  Construct 95% confidence intervals for the body mass of penguins (*body_mass_g*) for each species of penguin (*species*).

```{r}
require(dplyr)
alpha <- 0.05

penguins %>% 
    group_by(species) %>% 
    summarize(lower = mean(body_mass_g, na.rm=TRUE) - qt(1- alpha/2, (n() - 1))*sd(body_mass_g, na.rm=TRUE)/sqrt(n()),
              upper = mean(body_mass_g, na.rm=TRUE) + qt(1- alpha/2, (n() - 1))*sd(body_mass_g, na.rm=TRUE)/sqrt(n()))
```

95% CI for Adelie species' body mass (in grams) is (3627.2, 3774.2) 
95% CI for Chinstrap species' body mass (in grams) is (3640.1, 3826.1) 
95% CI for Gentoo species' body mass (in grams) is (4986.4, 5165.6) [(these are the chonky bois)]{style="color: #a632a8"}

5.  Construct a scatterplot with body mass (*body_mass_g*) on the *y*-axis, flipper length (*flipper_length_mm*) on the *x*-axis, and the markers (dots) colored by species of penguin (*species*).

```{r}
penguins %>% 
   ggplot(aes(x = flipper_length_mm, color=species)) +
   geom_point(aes(y = body_mass_g)) +
   labs(y= "body mass (g)", x= "flipper length (mm)")
   theme_bw()
```

## Thursday, August 31

6.  Model penguin body mass (*body_mass_g*) as a function of flipper length (*flipper_length_mm*) and bill length (*bill_length_mm*). Remember to state the resulting regression model.

```{r}
model1 <- glm(body_mass_g ~ flipper_length_mm + bill_length_mm, 
          data = penguins,
          family = "gaussian")
 summary(model1)
```
Resulting regression model:
y_hat = -5736.90 + 48.15x1 + 6.05x2
where:
y_hat = body mass (g)
x1 = flipper length (mm)
x2 = bill length (mm)

7.  Create a table that has Estimate (95% CI) and *p*-Value as columns; every predictor in the regression model will have a row. See [Table 2 in this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4742426/) for an example.

```{r}
summary(model1, level=0.95)
```

8.  Use the appropriate hypothesis test to determine if this is a significant regression line. Test at the $\alpha=0.05$ level. Remember to typeset your results.

```{r}
full <- glm(body_mass_g ~ flipper_length_mm + bill_length_mm, data = penguins, family = "gaussian")
reduced <- glm(body_mass_g ~ 1, data = penguins, family = "gaussian")
anova(reduced, full, test = "F")
```
For full model:
Reject Ho. There is sufficient evidence to suggest that the slope is non-zero.

9.  Use the appropriate hypothesis test(s) to determine which, if any, are significant predictors of penguin body mass. Test at the $\alpha=0.05$ level. Remember to typeset your results.

```{r}
summary(model1)
```
For flipper length:
Reject Ho. There is sufficient evidence to suggest that flipper length significantly predicts body mass.

For bill length:
Fail to reject Ho. There is not sufficient evidence that bill length significantly predicts body mass.

10. Create a data visualization for your regression model. Use the plot in question 5 as a starting point and overlay a regression line. (This means that only flipper length will vary and you will need to hold bill length constant.)

```{r}
coef <- coefficients(model1)
penguins %>%
summarize(mean=mean(bill_length_mm, na.rm=TRUE))
          
penguins <- penguins %>%
  mutate(y_hat_flipper = coef["(Intercept)"] + coef["flipper_length_mm"]*flipper_length_mm + coef["bill_length_mm"]*44)

penguins %>% 
   ggplot(aes(x = flipper_length_mm, color=species)) +
   geom_point(aes(y = body_mass_g)) +
   geom_line(aes(y = y_hat_flipper), color = "black") +
   geom_text(aes(x = 176, y = 2700, label = "bill length = avg"), color = "black", size = 4)
   labs(y= "body mass (g)", x= "flipper length (mm)") +
   theme_bw()
```

```

5. Construct a scatterplot with body mass (*body_mass_g*) on the *y*-axis, flipper length (*flipper_length_mm*) on the *x*-axis, and the markers (dots) colored by species of penguin (*species*).

```{r}

```

## Thursday, August 31

6. Model penguin body mass (*body_mass_g*) as a function of flipper length (*flipper_length_mm*) and bill length (*bill_length_mm*). Remember to state the resulting regression model.

```{r}

```

7. Create a table that has Estimate (95% CI) and *p*-Value as columns; every predictor in the regression model will have a row. See [Table 2 in this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4742426/) for an example.

```{r}

```

8. Use the appropriate hypothesis test to determine if this is a significant regression line. Test at the $\alpha=0.05$ level. Remember to typeset your results.

```{r}

```

9. Use the appropriate hypothesis test(s) to determine which, if any, are significant predictors of penguin body mass. Test at the $\alpha=0.05$ level. Remember to typeset your results.

10. Create a data visualization for your regression model. Use the plot in question 5 as a starting point and overlay a regression line. (This means that only flipper length will vary and you will need to hold bill length constant.)

```{r}

```


















